---
title: "Coronavirus misinformation analysis"
author: "Daniel Allington"
date: "2024-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(readr)
library(ggplot2)

options(scipen = 999)

misinfo_d <- read_csv('../data/life-under-lockdown-wave2-extract.csv')
```

# Association between misinformation beliefs and gender
## Hypothesis
It was hypothesised that number of misinformation beliefs would be associated with gender.

## Methodology
Mean numbers of beliefs for male and female respondents were compared using the Welch two-sample t-test. Respondents who did not identify as male or female were excluded from this analysis.

## Findings
```{r}
misinfo_d %>%
  ggplot(aes(x = gender)) +
  geom_bar()
```

```{r}
misinfo_d %>%
  ggplot(aes(x = beliefs)) +
  geom_histogram(bins = 6)
```

```{r}
mean(misinfo_d$beliefs)
```

```{r}
misinfo_d %>%
  filter(gender %in% c('Male', 'Female')) %>%
  t.test(formula = beliefs ~ gender, data = .)
```

In the sample, males had more misinformation beliefs (M = 0.73) than females (M = 0.61), but this was not statistically significant, t(491.31) = -1.201, p = 0.230.

# Association between misinformation beliefs and age
## Hypothesis
It was hypothesised that there would be a positive correlation between age and number of misinformation beliefs.

## Methodology
The product-moment or Pearson coefficient of correlation between age and number of misinformation beliefs was calculated.

## Findings
```{r}
misinfo_d %>%
  ggplot(aes(x = age, y = beliefs)) +
  geom_jitter()
```

```{r}
misinfo_d %>%
  cor.test(formula = ~ age + beliefs, data = ., method = 'pearson')
```

There was a statistically significant moderate negative correlation between age and number of misinformation beliefs, r(498) = -.24, p < .001. This appears to be accounted for by the fact that the minority of people who believe in many items of misinformation are mostly young. However, across all ages, most respondents had few or no misinformation beliefs.

# Association between misinformation beliefs and amount of knowledge about coronavirus which comes from Facebook

## Hypothesis
It was hypothesised that there would be a positive correlation between amount of knowledge which comes from Facebook and number of misinformation beliefs.

## Methodology
The rank-order or Spearman coefficient of correlation between amount of knowledge which comes from Facebook and number of misinformation beliefs was calculated.

## Findings
```{r}
misinfo_d <- misinfo_d %>%
  mutate(
    know_facebook_coded = know_facebook %>%
      recode(
        "Nothing at all" = 0,
        "Not very much" = 1,
        "A fair amount" = 2,
        "A great deal" = 3
      )
  )

misinfo_d %>%
  cor.test(formula = ~ know_facebook_coded + beliefs, data = ., method = 'spearman')
```

There was a statistically significant moderate positive correlation between number of misinformation beliefs and amount of knowledge which comes from Facebook, rs(498) = .22, p < .001.

# Association between age and amount of knowledge about the coronavirus which comes from Facebook
```{r}
misinfo_d %>%
  cor.test(formula = ~ know_facebook_coded + age, data = ., method = 'spearman')
```

# Association between age and number of misinformation beliefs about the coronavirus, controlling for amount of knowledge about the coronavirus which comes from Facebook
```{r}
misinfo_d <- misinfo_d %>%
  mutate(
    know_facebook_binary = 
      as.numeric(know_facebook == 'A great deal'))

misinfo_d %>%
  lm(
    formula = beliefs ~ age + know_facebook_binary, data = .
  ) %>%
  summary
```
Number of misinformation beliefs is predicted by age and amount of knowledge from Facebook, F(2,497) = 22.56, p < .001, r2 = .08. Both predictors were statistically significant, p < .001.


